\section{Complex Structures}

Let $X$ be a topological space.

\begin{boxdefinition}[$\Delta$-complex]
    A $\Delta$-complex structure on $X$ is a set of singular $n$-simplices $\delta_{\alpha} : '\Delta^n \to X$ for varying $n = n(\alpha)$ such that
    \begin{enumerate}
        \item every $x \in X$ is the image of at least one $\delta_{\alpha}$, $\sigma_{\alpha} \vert_{\interior{\Delta^n}}$ is injective
        \item $\sigma_{\alpha}\vert_{\brac{v_0, \ldots, \widehat{v_i}, \ldots, v_n}}$ is $\sigma_{\beta} : \Delta^{k-1} \to X$ in the order-preserving identification of vertices
        \item $U \ssq X$ is open iff $\sigma_{\alpha}\inv\of{U}$ is open for all $\alpha$.
    \end{enumerate}
\end{boxdefinition}

This is the ``minimal definition'' for making the definition of homology work.

Our objective will be to show that homotopies of spaces induce homotopies of chain complexes.

\subsection{Prisms}

\begin{boxdefinition}[$n$-Prism]
    For all $n \in \N$, define the $n$-prism to be $\Delta^n \times I \ssq \R^n$.
\end{boxdefinition}

This is not a simplex, so we need to triangulate: we can view the $n$-prism as a(n uncountable) union of $(n-1)$-simplices. Denote the bottom $n$-simplex of an $n$-prism by $v_0, \ldots, v_n$ and the top $n$-simplex by $w_0, \ldots, w_n$.

\begin{boxdefinition}[Prism Operator]
    Define $P : C_n(X) \to C_{n+1}(X)$ by
    \begin{align*}
        P(\sigma) = \sum_{i} (-1)^{i} F \circ (\sigma \times \id) \vert_{\brac{v_0, \ldots, v_i, w_i, \ldots, w_n}}
    \end{align*}
\end{boxdefinition}

\begin{boxproposition}
    Given $f, g : X \to Y$, if
    \begin{align*}
        \del \circ P = g_{\#} - f_{\#} - P \circ \del
    \end{align*}
    then the maps induced by $f$ and $g$ on homology
    \begin{align*}
        f_*, g_* : H_*(X) \to H_*(Y)
    \end{align*}
    are equal.
\end{boxproposition}

\begin{boxdefinition}[Chain Homotopy]
    A sequence of maps $P_n : C_n(X) \to C_{n+1}(X)$ is called a \textbf{chain homotopy} if
    \begin{align*}
        \del \circ P + P \circ \del = g_{\#} - f_{\#}
    \end{align*}
\end{boxdefinition}

We keep in mind the following commutative diagram:
\begin{cd*}
    \cdots &[2em] {C_{n+1}(X)} &[4em] {C_n(X)} &[4em] {C_{n-1}(X)} &[2em] \cdots \\[4em]
	\cdots &[2em] {C_{n+1}(X)} &[4em] {C_n(Y)} &[4em] {C_{n-1}(X)} &[2em] \cdots
	\arrow["\partial", from=1-1, to=1-2]
	\arrow["\partial", from=1-2, to=1-3]
	\arrow["{f_{\#}}"', shift right, from=1-2, to=2-2]
	\arrow["{g_{\#}}", shift left, from=1-2, to=2-2]
	\arrow["\partial", from=1-3, to=1-4]
	\arrow["P"{description}, from=1-3, to=2-2]
	\arrow["{f_{\#}}"', shift right, from=1-3, to=2-3]
	\arrow["{g_{\#}}", shift left, from=1-3, to=2-3]
	\arrow["\partial", from=1-4, to=1-5]
	\arrow["P"{description}, from=1-4, to=2-3]
	\arrow["{g_{\#}}", shift left, from=1-4, to=2-4]
	\arrow["{f_{\#}}"', shift right, from=1-4, to=2-4]
	\arrow["\partial"', from=2-1, to=2-2]
	\arrow["\partial"', from=2-2, to=2-3]
	\arrow["\partial"', from=2-3, to=2-4]
	\arrow["\partial"', from=2-4, to=2-5]
\end{cd*}

\subsection{Exact Sequences}

Let $A \ssq X$. By $C_n(A)$, we denote the group $C_n(A; \Z)$ - the free abelian group generated by continuous functions from $\Delta^n$ to $A$.

It is, of course, a subgroup of $C_n(X)$, and since everything is abelian, all subgroups are normal, so we can construct a chain complex of quotients
\begin{align*}
    C_n\of{X, A} := \quotient{C_n(X)}{C_n(A)}
\end{align*}
\begin{boxtheorem}[Zigzag Theorem/Long Exact Sequence of Homology]\label{SP:Thm:LES_Homology}
    A short exact sequence $0\rightarrow A_{*}\rightarrow B_{*}\rightarrow C_{*} \to 0$ of chain complexes induces a long exact sequence of homology groups.
    % \begin{cd*}
    %     \cdots 
    %     \arrow[r] 
    %     & H_n(A) 
    %     \arrow[r] 
    %     & H_n(B) 
    %     \arrow[r] 
    %     & H_n(C) 
    %     \arrow[dll, 
    %          decorate,
    %          decoration={snake, amplitude=1.2pt, segment length=6pt},
    %          "\delta_n"'] \\
    %     H_{n-1}(A) 
    %     \arrow[r] 
    %     & H_{n-1}(B) 
    %     \arrow[r] 
    %     & H_{n-1}(C) 
    %     \arrow[dll, 
    %          decorate,
    %          decoration={snake, amplitude=1.2pt, segment length=6pt},
    %          "\delta_{n-1}"'] \\
    %     H_{n-2}(A) 
    %     \arrow[r] 
    %     & H_{n-2}(B) 
    %     \arrow[r] 
    %     & H_{n-2}(C) 
    %     \arrow[r] 
    %     & \cdots
    % \end{cd*}
\end{boxtheorem}
\begin{proof}
    See \cite[Week 6, Lemma 1.6]{ICLAlg4} for a proof that uses the Snake Lemma or \cite[Proposition 5.5.1]{EPFLRingsAndModules} for a more fleshed out version.
\end{proof}

Recall the following definition.

\begin{boxdefinition}
    For $A \ssq X$, we denote by $C_{\bullet}(X, A)$ the chain complex given by
    \begin{align*}
        C_n(X, A) = \quotient{C_n(X)}{C_n(A)}
    \end{align*}
    for all $n \in \N$.
\end{boxdefinition}

\begin{boxtheorem}[Long Exact Sequence of a Pair]
    For all $A \ssq X$, there is a LES
    \begin{cd*}
        \cdots \arrow[r] & H_n(A) \arrow[r] & H_n(X) \arrow[r] & H_n(X, A) \arrow[r] & H_{n-1}(A) \arrow[r] & \cdots
    \end{cd*}
\end{boxtheorem}
\begin{proof}
    Exactly \Cref{SP:Thm:LES_Homology} applied to the SES
    \begin{cd*}
        0 \arrow[r] & C_{\bullet}\of{A} \arrow[r] & C_{\bullet}\of{X} \arrow[r] & C_{\bullet}\of{X, A} \arrow[r] & 0
    \end{cd*}
\end{proof}
Say now that we have a space $X=A\cup B$ for some open $A,B$. We would like to say that $0\rightarrow C_n(A\cap B)\rightarrow C_n(A)\oplus C_n(B)\rightarrow C_n(X)\rightarrow 0$ is short exact, but this is not quite true - we can imagine a simplex $S$ which is only partially contained in each of $A,B$. However there is an easy fix for this (and it is what you would expect).

\begin{boxnotation}
    Let $X$ be a space that is covered by open sets $A$ and $B$. Define
    \begin{align*}
        C^{\set{A, B}}_n(X)
    \end{align*}
    to be the subgroup of $C_n(X)$ generated by those $\sigma : \Delta^n \to X$ such that $\pim{\sigma} \ssq A$ or $\pim{\sigma} \ssq B$.
\end{boxnotation}

The idea of the above is that we are somehow ``splitting simplices into two parts''---the part contained in $A$ and the part conained in $B$.

\begin{boxtheorem}
    Given a space $X$ and open sets $A, B \ssq X$ such that $X = A \cup B$, the following is a short exact sequence of chain complexes:
    \begin{cd*}
        0 \arrow[r] & C_{*}\of{A \cap B} \arrow[r] & C_{*}(A) \+ C_{*}(B) \arrow[r] & C^{\set{A, B}}_{*}(X) \arrow[r] & 0
    \end{cd*}
\end{boxtheorem}

Note that in the above SES, if we were to replace $C_{*}^{\set{A, B}}(X)$ by $C_{*}(X)$, the sequence would cease to be exact.

\begin{boxlemma}
    For all $n \in \N$, the $n$th homology group of $C_{*}(X)$ is the same as the $n$th homology group of $C_{*}^{\set{A, B}}(X)$.
\end{boxlemma}

More generally, we show the following.

\begin{boxlemma}
    Let $\U = \setst{U_j \ssq X}{j < J}$ be subsets of $X$ whose interiors cover $X$. Define
    \begin{align*}
        C_n^{\U}(X) := \setst{\sum_i \lambda_i \sigma_i \in C_n(X)}{\forall i,\ \exists j \st \pim{\sigma_i} \ssq U_j}
    \end{align*}
    The inclusions $i_{*} : C_{*}^{\U}(X) \inj C_{*}(X)$ form a homotopy equivalence of chain complexes.
\end{boxlemma}

By the Zigzag theorem, this gives an LES in homology:
\begin{boxdefinition}[Mayer-Vietoris Sequence]
    Given $X$ covered by the interiors of two sets $A,B$ we call the LES
    \begin{cd*}
        \cdots \arrow[r]
        &[0.1em] H_n\of{A \cap B} \arrow[r]
        & H_n(A) \+ H_n(B) \arrow[r]
        & H_n(X) \arrow[r]
        & H_{n-1}\of{A \cap B} \arrow[r]
        % & H_{n-1}(A) \+ H_{n-1}(B) \arrow[r]
        &[0.1em] \cdots
    \end{cd*}
    the \textbf{Mayer-Vietoris Sequence for $X$}.
\end{boxdefinition}

Say we have a space $X$ and $\mathcal{U}=\{U_j:j\in J\}$ is a cover of $X$ and the interiors of the $U_j$ also cover $X$. We let $C_n^{\mathcal{U}}\ssq C_n(X)$ consist of all chains $\sum_{i\in I} \lambda_i \sigma_i$ where for all $i\in I$ we have $\text{im}(\sigma_i)\ssq U_j$ for some $j$. Note that the boundary maps $\partial$ restrict to boundary maps $\partial: C_n{\mathcal{U}}\rightarrow C_{n-1}^{\mathcal{U}}(X)$. 

We now temporarily call the corresponding homology groups $H_n^{\mathcal{U}}(X)$ (``temporarily'' because we will hopefully soon show that $H_n^{\mathcal{U}}(X)$ is just $H_n(X)$ for every $\mathcal{U}$):
\begin{boxlemma}
    The inclusion map $i:C_n^{\mathcal{U}}(X)\rightarrow C_n(X)$ is a chain homotopy equivalence (i.e. the projection $p:C_n(X)\rightarrow C_n^{\mathcal{U}}(X)$ is such that $p\circ i$ and $i\circ p$ are chain homotopic to the identity), and in particular we have $H_n^{\mathcal{U}}(X)\cong H_n(X)$.
\end{boxlemma}
\begin{proof}
    We just split $\Delta^n$ into many smaller simplices, for example by repeated barycentric subdivisions... we then find a homotopy of chain complexes between $C_{*}(X)$ and $C_{*}^{\set{A, B}}(X)$.

    Let $Y \ssq \R^d$ be a convex set, and let $LC_n(Y) \ssq C_n(Y)$ denote the subgroup of affine maps $\sigma : \Delta^n \to Y$. The boundary map $\partial$ restricts to a map $\partial : LC_n(Y) \to LC_{n-1}(Y)$. For every $b \in Y$, define
    \begin{align*}
        b : LC_n(Y) \to LC_{n+1}(Y) : \brac{w_0, \dots, w_n} \mapsto \brac{b, w_0, \ldots, w_n}
    \end{align*}
    It is not difficult to show that
    \begin{align*}
        \partial \circ b = \id - b \circ \partial \iff \partial \circ b + b \circ \partial = \id
    \end{align*}

    We now let $S:LC_n(Y)\rightarrow LC_n(Y)$ be defined inductively by $S:LC_0(Y)\rightarrow LC_0(Y)$ is the identity, and $S_{\sigma}=b_{\sigma}(S\partial \sigma)$, where $B_{\sigma}$ is the barycenter of $S\partial \sigma$. Then $\partial S_{\sigma}=\partial b_{\sigma}(S\partial \sigma)=S\partial \sigma -b_{\sigma}\partial $\sorry
\end{proof}

\subsection{Application: Homology Groups of Spheres}

Let $B_n$ denote the closed unit ball in $\R^n$ centred at the origin, so that $S^{n-1} = \partial B_n \subseteq B_n$. Then, the following is a long exact sequence:
\begin{cd*}
    \cdots \arrow[r] & H_k\of{S^{n-1}} \arrow[r] & H_k\of{B_n} \arrow[r] & H_k\of{B_n, S^{n-1}} \arrow[r] & H_{k-1}\of{S^{n-1}} \arrow[r] & \cdots
\end{cd*}

We can show the following fact.

\begin{boxtheorem}
    $H_n\of{S^n; \Z} \cong \Z$ for all $n \geq 1$.
\end{boxtheorem}
\begin{proof}
    
\end{proof}

% Next up: Mayer-Vietoris

\subsection{Excision}

(Recall that relative homology is obtained by quotienting out chain complexes in $X$ by chain complexes in $A$ and then computing homology.)

\begin{boxtheorem}[Excision]
    Let $A, B \ssq X$, with the interiors covering $X$. The inclusion of $(B, A \cap B)$ into $(X, A)$ induces isomorphisms
    \begin{align*}
        H_ n\of{B, A \cap B} \xrightarrow{\sim} H_n\of{X, A}
    \end{align*}
    for all $n \in \N$. Equivalently, suppose we have a chain of inclusions $Z \ssq A \ssq X$, with $\cl{Z} \ssq \interior{A}$. Then,
    \begin{align*}
        H_n(X, A) \cong H_n\of{X \setminus Z, A \setminus Z}
    \end{align*}
    These equivalent phenomena are called \textbf{excision}.
\end{boxtheorem}
\begin{proof}
    We have a diagram
    \begin{cd*}
        \quotient{C_n(B)}{C_n(A \cap B)} \arrow[r] &
        \quotient{C_n^{\set{A, B}}(X)}{C_n(A)} \arrow[r, "\text{induces isomorphism on $H_*$}"] &[8em]
        \quotient{C_n(X)}{C_n(A)}
    \end{cd*}
\end{proof}

We can also talk about nested inclusions.

\begin{boxproposition}[Long Exact Sequence of a Triple]
    Given inclusions $A \ssq B \ssq X$, the following is a short exact sequence of chain complexes:
    \begin{cd*}
        0 \arrow[r] & C_{\bullet}\of{B, A} \arrow[r] & C_{\bullet}\of{X, A} \arrow[r] & C_{\bullet}\of{X, B} \arrow[r] & 0
    \end{cd*}
\end{boxproposition}

This again induces a long exact sequence of homology.

\begin{boxexample}[$\star \ssq S^1 \ssq D^2$]
    Consider a point $\star \in S^1$ and consider the inclusions given by $\set{\star} \ssq S^1 \ssq D^2$. We get a long exact sequence
    \begin{cd*} % TODO: Fix formatting
        \cdots \arrow[r] &[0.25em]
        H_2\of{S^1, \star} \arrow[r] &[0.25em]
        H_2\of{D^2, \star} \arrow[r] &[0.25em]
        H_2\of{D^2, S^1} \arrow[r] &[0.25em]
        H_1\of{S^1, \star} \arrow[r] &[0.25em]
        H_1\of{D^2, \star} \arrow[r] &[0.25em]
        H_1\of{D^2, S^1} \arrow[r] &[0.25em]
        H_0\of{S^1, \star} \arrow[r] &[0.25em]
        \cdots
    \end{cd*}
    We can use what we know about long exact sequences of a pair to compute $H_n\of{S^1, \star}$:
    \begin{cd*} % TODO: Fix formatting
        \cdots \arrow[r] &[0.25em]
        \underbrace{H_n\of{\star}}_{=0} \arrow[r] &[0.25em]
        H_n\of{S^1} \arrow[r, "\sim"] &[0.25em]
        H_n\of{S^1, \star} \arrow[r] &[0.25em]
        \underbrace{H_{n-1}\of{\star}}_{=0} \arrow[r] &[0.25em]
        \cdots
    \end{cd*}
    In particular, $H_n\of{S^1, \star} \cong \Tilde{H_n}\of{S^1, \star}$. (And the same is true of other spaces as well, not just $S^1$.)
    % This tells us what? How?
\end{boxexample}

\subsection{Changing the Base Group}

Suppose I have a short exact sequence of abelian groups
\begin{cd*}
    0 \arrow[r] & A' \arrow[r] & A \arrow[r] & A'' \arrow[r] & 0
\end{cd*}

It turns out this induces the following long exact sequence of chain complexes:
\begin{cd*}
    0 \arrow[r] & C_{\bullet}\of{X; A'} \arrow[r] & C_{\bullet}\of{X; A} \arrow[r] & C_{\bullet}\of{X; A''} \arrow[r] & 0
\end{cd*}
where at every stage, we compute the cycle groups with respect to a different base group.

\subsection{Good Pair}

\begin{boxdefinition}[Good Pair]
    A pair $(X, A)$ of some $A \ssq X$ is a \textbf{good pair} if $A$ is nonempty and closed and there exists a neighbourhood $V$ of $A$ in $X$ that deformation retracts to $A$.
\end{boxdefinition}

% Something something big pair property

\subsection{Reconciling Singular and Simplicial Homology}

Recall the Five Lemma.

\begin{boxlemma}[Five Lemma]
    Consider the following diagram in some abelian category, with the top and bottom rows both being exact:
    \begin{cd*}
        A & B & C & D & E \\
    	{A'} & {B'} & {C'} & {D'} & {E'}
    	\arrow["i", from=1-1, to=1-2]
    	\arrow["\alpha"{description}, from=1-1, to=2-1]
    	\arrow["j", from=1-2, to=1-3]
    	\arrow["\beta"{description}, from=1-2, to=2-2]
    	\arrow["k", from=1-3, to=1-4]
    	\arrow["\gamma"{description}, from=1-3, to=2-3]
    	\arrow["l", from=1-4, to=1-5]
    	\arrow["\delta"{description}, from=1-4, to=2-4]
    	\arrow["\epsilon"{description}, from=1-5, to=2-5]
    	\arrow["{i'}", from=2-1, to=2-2]
    	\arrow["{j'}", from=2-2, to=2-3]
    	\arrow["{k'}", from=2-3, to=2-4]
    	\arrow["{l'}", from=2-4, to=2-5]
    \end{cd*}
    The following are all true.
    \begin{enumerate}
        \item If $\beta, \delta$ are surjective and $\epsilon$ is injective, then $\gamma$ is surjective.
        \item If $\beta, \delta$ are injective and $\alpha$ is surjective, then $\gamma$ is injective.
    \end{enumerate}
    In particular, if $\alpha, \beta\, \delta, \epsilon$ are isomorphisms, then $\gamma$ is also an isomorphism.
\end{boxlemma}
\begin{proof}
    We begin the grand diagram chase. Note: we can also prove the first two as Four Lemmas - see \cite[Problem Sheet 5, Exercise 2]{EPFLRingsAndModules}. For a version of this statement, see \cite[Lemma 5.6.2]{EPFLRingsAndModules}.
    \begin{enumerate}
        \item Fix $c' \in C'$. There is $d \in D$ such that $\delta(d) = k'\of{c'}$. Notice that $\epsilon\of{l(d)} = l'\of{\delta(d)} = l'\of{k'\of{c'}} = 0$. Since $\epsilon$ is injective, this tells us that $l(d) = 0$. Thus, $d = k(c)$ for some $c \in C$. Then,
        \begin{align*}
            k'\of{c' - \gamma(c)}
            &= k'\of{c'} - k'\of{\gamma(c)} \\
            &= k'\of{c'} - \delta\of{k(c)} \\
            &= k'\of{c'} - \delta(d)
            &= 0
        \end{align*}
        \sorry

        \item Fix $c \in C$ with $\gamma(c)$
    \end{enumerate}
    The union of the first two gives you immediately that if $\alpha$, $\beta$, $\delta$ and $\epsilon$ are isomorphisms, then so is $\gamma$.
\end{proof}

We are now ready to show that singular and simplicial homology are isomorphic. In fact, they are naturally isomorphic as functors.

\begin{boxtheorem}
    Let $X$ be a $\Delta$-complex and let $A \ssq X$ be a subcomplex. Then,
    \begin{align*}
        \underbrace{H_{\bullet}^{\Delta}\of{X, A}}_{\text{simplicial}}
        \cong
        \underbrace{H_{\bullet}\of{X, A}}_{\text{singular}}
    \end{align*}
    naturally in $X$ and $A$.
\end{boxtheorem}
\begin{proof}
    
\end{proof}

\begin{boxexample}[A Weird Pathology]
    Define a $\Delta$-complex by gluing together countably many copies of the unit interval at the same endpoint. So it looks like an asterisk with countably many spokes (the caveat being this space is actually not metrisable, and in particular, doesn't embed into any Euclidean space, let alone the plane). A path that goes halfway along the first spoke before going back to the vertex, a quarter of the way along the second before going back to the vertex, an eight of the way along the third before going back, and so on. 
\end{boxexample}