\section{The Setup}

Before we go any further, we will state something (that is par for any course on topology).

\begin{boxconvention}
    Unless otherwise specified,
    \begin{itemize}
        \item All maps in this course will be assumed to be continuous.
        \item The (uppercase) letters $X$, $Y$ and $Z$ will denote topological spaces.
    \end{itemize}
\end{boxconvention}

\subsection{Paths and Homotopies}

We begin by defining the notion of `deformation' we talked about so handwavily just a few moments ago.

\begin{boxdefinition}[Homotopy of Continuous Maps]
    Let $X$ and $Y$ be topological spaces and let $f, g : X \to Y$ be (continuous) functions. A \textbf{homotopy from $f$ to $g$} is a (continuous) map $H : X \times I \to Y$ such that for all $x \in X$,
    \begin{align*}
        H(x, 0) = f(x) \qquad \text{and} \qquad H(x, 1) = g(x)
    \end{align*}
    If a homotopy exists between $f$ and $g$, we say they are \textbf{homotopic}, and we write $f \simeq g$.
\end{boxdefinition}

It is not hard to see that homotopy (or more precisely, the property of being homotopic) is an equivalence relation on the set of continuous functions from one space to another. In particular, it doesn't matter whether we say `$f$ is homotopic to $g$' or `$g$ is homotopic to $f$'.

\begin{boxdefinition}[Paths and Loops]
    A \textbf{path} in a topological space $X$ is a continuous map $\gamma : [0, 1] \to X$. A path $\gamma$ is a \textbf{loop} if $\gamma(0) = \gamma(1)$.
\end{boxdefinition}

Since life is too short, we will adopt the following notation.

\begin{boxnotation}
    Unless otherwise specified, the symbol $I$ will denote the unit interval $[0, 1]$.
\end{boxnotation}

There are obviously many examples of paths. In fact, there are many examples of homotopic paths.

\begin{boxexample}
    Any two paths in $\R^n$ (with the same endpoints) are homotopic. For example, if $\alpha, \beta : I \to \R^n$ are paths with the same endpoints, then
    \begin{align*}
        H : I \times I \to \R^n : (s, t) \mapsto t \cdot \alpha(s) + (1 - t) \cdot \beta(s)
    \end{align*}
    is a homotopy between the paths.
\end{boxexample}

Really, it only makes sense to talk about homotopies between paths whose endpoints are the same. That being said, our definition of homotopies is permissive enough that this is not necessarily true of a homotopy. Therefore, we will introduce separate notation for what it means for paths with the same endpoints to be homotopic (notation that emphasises that their endpoints are the same).

\begin{boxnotation}
    If $\alpha$ and $\beta$ with the same endpoints, we will write
    \begin{align*}
        \alpha \hdi \beta
    \end{align*}
    to indicate that $\alpha$ and $\beta$ are homotopic.
\end{boxnotation}

It turns out that paths can be composed, provided they are `compatible'.

\begin{proposition}[Composing Paths]
    Fix points $x_0, x_1, x_2 \in X$. If $\gamma$ is a path from $x_0$ to $x_1$ and $\delta$ is a path from $x_1$ to $x_2$, then the function
    \begin{align*}
        \eps : [0, 1] \to X : t \mapsto
        \begin{cases}
            \gamma\of{2t} & \text{ if } 0 \leq t \leq \frac{1}{2} \\
            \delta\of{2t - 1} & \text{ if } \frac{1}{2} \leq t \leq 1
        \end{cases}
    \end{align*}
    is a path between $x_0$ and $x_2$.
\end{proposition}
\begin{proof}
    It is clear that $\eps$ is a well-defined function: at the point $t = \frac{1}{2}$,
    \begin{align*}
        \gamma\of{2t} = \gamma\of{1} = x_1 = \delta\of{0} = \delta\of{2t - 1}
    \end{align*}
    So indeed the function is well-defined.

    All we need to show is that $\eps$ is continuous. But this is true because of a homework exercise from General Topology\todo{Add ref} in which we showed that if we have a space that is a union of two closed sets then we can continuous `glue' together continuous functions that agree on the overlap of these closed sets. So $\eps$ is the `gluing' of $\gamma$ of $\delta$, and is hence continuous.
\end{proof}

\begin{boxdefinition}[Composition of Paths]
    Fix points $x_0, x_1, x_2 \in X$. If $\gamma$ is a path from $x_0$ to $x_1$ and $\delta$ is a path from $x_1$ to $x_2$, then the function $\eps : [0, 1] \to X$ defined as above is called the \textbf{composition} of $\gamma$ and $\delta$.
    
    We will denote it $\delta \star \gamma$, $\delta * \gamma$, $\delta \cdot \gamma$ or simply $\delta \gamma$.
\end{boxdefinition}

\begin{boxwarning}
    We emphasise that $\delta \gamma$ means
    \begin{center}
        \textbf{First do $\gamma$, then do $\delta$.}
    \end{center}
    In other words, we obey the convention we normally use when we do function composition.
\end{boxwarning}

\begin{boxwarning}
    You know everything I just said? Be warned, I may not always heed my own warning and might write $\gamma \delta$ to mean $\delta \gamma$ sometimes. The real warning is, just use your head, typecheck that function compositions and path concatenations are all ok, and you should be fine. Just read the vibe and you should be ok.
\end{boxwarning}

Finally, we define the notion of an inverse path.

\begin{boxdefinition}[Inverse Path]
    Given a path $\alpha : I \to X$, we define the \textbf{inverse path} $\bar{\alpha} : I \to X$ by $\bar{\alpha}(t) := \alpha(1 - t)$ for all $t \in I$.
\end{boxdefinition}

We will now take a brief detour into the world of groupoids.

\subsection{Groupoids}

Here, we will assume some familiarity with Category Theory. For a quick and dirty introduction to the subject, read \sorry.\todo{Add refs}

Recall that we can view a group as a (small) category, with the morphisms being its action on itself.

\begin{boxexample}[A Group as a (Small) Category]\label{Ch1:Eg:Group_as_Category}
    Fix a group $G$. We can turn it into a small category with one single object $\star$ and one morphism from $\star$ to itself corresponding to each group element $g$.
\end{boxexample}

We define a groupoid to be a generalisation of the above example that allows for more than one object.

\begin{boxdefinition}[Groupoid]\label{Ch1:Def:Groupoid}
    A \textbf{groupoid} is a small category in which every morphism is invertible.
\end{boxdefinition}

Comparing \Cref{Ch1:Def:Groupoid} and \Cref{Ch1:Eg:Group_as_Category}, the following is evident.

\begin{boxproposition}
    A group is a groupoid with only one object.
\end{boxproposition}

\subsection{The Fundamental Group(oid)}

Fix a topological space $X$. We ask ourselves the following question: can we associate a groupoid to $X$? As it turns out, we can.

\begin{boxdefinition}
    The \textbf{fundamental groupoid of $X$}, denoted $\pi_1(X)$, is the category with the following data:
    \begin{itemize}
        \item The object set is $X$.
        \item For each $x, y \in X$, define the set of morphisms $\Hom(x, y)$ to be the set of all paths from $x$ to $y$, up to homotopy.
        \item For $x, y, z \in X$, $[\gamma] \in \Hom(x, y)$ and $[\delta] \in \Hom(y, z)$, define $[\delta] \circ [\gamma]$ to be $[\delta \star \gamma]$.
        \item For $x, y \in X$ and $[\gamma] \in \Hom(x, y)$, define the inverse operation $[\gamma]\inv$ to be $\brac{\bar{\gamma}}$.
    \end{itemize}
\end{boxdefinition}

We have not really shown that this composition is associative (and of course it is not something that should be shown \textit{after} the definition but rather something that should be shown \textit{before}) but we'll do it now.

\begin{boxlemma}
    Let $\phi : I \to I$ be a path such that $\phi(0) = 0$ and $\phi(1) = 1$. For any path $\gamma : I \to X$, $\gamma \circ \phi \hdi \gamma$.
\end{boxlemma}
\begin{proof}
    The function
    \begin{align*}
        H : I \times I \to X : (s, t) \mapsto \gamma\of{st + (1-t)\phi(s)}
    \end{align*}
    is an explicit homotopy.
\end{proof}

\begin{boxcorollary}\label{Ch1:Cor:path_comp_assoc}
    For all paths $\alpha, \beta, \gamma : I \to X$ that are compatible,
    \begin{align*}
        \gamma \star \parenth{\beta \star \alpha} \hdi \parenth{\gamma \star \beta} \star \alpha
    \end{align*}
\end{boxcorollary}

Thus, concatenation gives rise to a well-defined (associative) composition on homotopy classes, thus, on the fundamental groupoid.

Recall that a group is a groupoid with a single object. So, it stands to reason that the fundamental groupoid, except when we look at just one point in our space (and paths from that point to itself) rather than \textit{all} points in our space (and \textit{all} paths), should be a group. This is precisely the fundamental group.

\begin{boxdefinition}[The Fundamental Group]
    Fix a point $x_0 \in X$. The \textbf{fundamental group of $X$ based at $x_0$} is the set
    \begin{align*}
        \pi_1\of{X; x_0} = \Hom_{\pi_1(X)}\of{x_0, x_0}
    \end{align*}
\end{boxdefinition}

Some fundamental groups are easy to compute and some are difficult.

\begin{boxexample}
    $\pi_1\of{\R^n}$ is trivial.
\end{boxexample}

\subsection{Basic Properties of the Fundamental Group}

\begin{boxtheorem}
    For any space $X$ and basepoint $x_0 \in X$, $\pi_1\of{X, x_0}$ is a group under the composition operation $\brac{\gamma} \cdot \brac{\gamma'} = \brac{\gamma \star \gamma'}$.
\end{boxtheorem}
\begin{proof}
    This should just follow trivially from the definition (because we did `check' that $\pi_1(X)$ is a groupoid), but let's play the game and check the axioms anyway.
    \begin{itemize}
        \item \underline{Identity:} The identity element is the loop $c_{x_0} : I \to X : t \mapsto x_0$. Indeed, $c_{x_0} \star \gamma \hdi \gamma \hdi \gamma \star c_{x_0}$ via
        \begin{align*}
            \Phi : I \to I : t \mapsto \begin{cases}
                2t & \text{ if } t \leq \frac{1}{2} \\
                1 & \text{ if } t \geq \frac{1}{2}
            \end{cases}
        \end{align*}

        \item \underline{Inverses:} Recall that we defined the inverse of a path $\gamma$ to be the path with the time parameter going in the opposite direction - $\bar{\gamma}(t) := \gamma(1 - t)$. It's not hard to show that for all loops $\gamma$ based at $x_0$, $\gamma \star \bar{\gamma} \hdi \bar{\gamma} \star \gamma$.

        \item \underline{Associativity:} This is precisely \Cref{Ch1:Cor:path_comp_assoc}.
    \end{itemize}
\end{proof}

\begin{boxtheorem}
    For $X$ a topological space with $x_0,x_1 \in X$, and $\alpha : I \to X$ a path from $x_0$ to $x_1$ ($\alpha(0)=x_0, \alpha(1)=x_1$), then $\pi_1(X,x_0) \cong \pi_1(X,x_1)$.
\end{boxtheorem}
\begin{quote}
    \textit{There is really only one reasonable way to attempt to do this. And when you do it, surprise surprise, it works...}
\end{quote}
\begin{proof}
    The desired isomorphism is given by conjugation by the path from $x_0$ to $x_1$. Explicitly, if $\alpha : I \to X$ is a path from $x_0$ to $x_1$, we can define $\Phi : \pi_1\of{X, x_0} \to \pi_1\of{X, x_1}$ by
    \begin{align*}
        \Phi\of{\brac{\gamma}} := \brac{\bar{\alpha} \star \gamma \star \alpha}
    \end{align*}
    It's not very hard to show that this is a group homomorphism: given paths $\brac{\gamma}, \brac{\gamma'} \in \pi_1\of{X, x_0}$, we see that
    \begin{align*}
        \Phi\of{\brac{\gamma} \circ \brac{\gamma'}}
        &= \brac{\bar{\alpha} \star \gamma \star \gamma' \star \alpha} \\
        &= \brac{\bar{\alpha} \star \gamma \star \alpha \star \bar{\alpha} \star \gamma' \star \alpha} \\
        &= \brac{\bar{\alpha} \star \gamma \star \alpha} \circ \brac{\bar{\alpha} \star \gamma' \star \alpha} \\
        &= \Phi\of{\brac{\gamma}} \circ \Phi\of{\brac{\gamma'}}
    \end{align*}
    Finally, we can define $\Phi\inv : \pi_1\of{X, x_1} \to \pi_1\of{X, x_0}$ by
    \begin{align*}
        \Phi\inv\of{\brac{\gamma}} := \brac{{\alpha} \star \gamma \star \bar{\alpha}}
    \end{align*}
    It's not hard to show this is a two-sided inverse.
\end{proof}

Finally, we show that $\pi_1$ is functorial.

\begin{boxproposition}
    Let $X$ and $Y$ be topological spaces. Fix a basepoint $x_0 \in X$. For any continuous function $f : X \to Y$, the map
    \begin{align*}
        f_* : \pi_1\of{X, x_0} \to \pi_1\of{Y, f\of{x_0}} : \gamma \mapsto \brac{f \circ \gamma}
    \end{align*}
    is a well-defined homomorphism of groups.
\end{boxproposition}
\begin{proof}
\sorry
    
\end{proof}

% Section break maybe? Let's see where this goes 

\begin{boxdefinition}[Homotopy Equivalence of Spaces] 
    Let $X$ and $Y$ be topological spaces. If there are maps $f : X \to Y$ and $g : Y \to X$ such that $f \circ g \simeq \id_Y$ and $g \circ f \simeq \id_X$, then we say $X$ and $Y$ are \textbf{homotopy equivalent}. We write $X \simeq Y$.
\end{boxdefinition}
% ^ Note this is trivially equal to saying categorical equivalence of the fundamental groupoid i think.

There are many well-known examples of homotopy equivalences of spaces. At some point, we will probably look at \textbf{contractible spaces}, which are spaces that are homotopic to a point.

\begin{boxexample}[Contractibility of the Unit Interval]
    $I \simeq \set{0}$ via the maps
    \begin{align*}
        f : I \to \set{0} : t \mapsto 0
        \qquad
        g : \set{0} \to I : 0 \mapsto 0
    \end{align*}
    Obviously, $f \circ g = \id_{\set{0}}$. We only need a homotopy in the other direction. Indeed, it's not hard to show that
    \begin{align*}
        H : I \times I \to I : (s, t) \mapsto st
    \end{align*}
    gives the homotopy $g \circ f \simeq \id_I$.
\end{boxexample}

We can see the above example as indicative of a more general pattern - if we have spaces $X$ and $Y$ with $Y\ssq X$, then it seems that $X$ and $Y$ should be homotopic if there are no ``holes'' in $X\setminus Y$.

\begin{boxdefinition}[Deformation Retraction]
    Let $X$ be a topological space. Fix $A \ssq X$ and let $i : A \inj X$ denote the inclusion. A map $r : X \to Y$ is a \textbf{deformation retraction} if $r \circ i = \id_A$ and $i \circ r \simeq \id_X$.
\end{boxdefinition}

See also \cite[Ch 4, \S 4, Définition 4.3]{EPFLTopologie}.

A special property of deformation retractions is that they induce inclusions of fundamental groups.

\begin{boxconvention}
    Going forward, unless otherwise specified, arbitrary spaces $X, Y, Z$ etc will be assumed to be path-connected (hopefully for obvious and unproblematic reasons). We will also often omit the basepoint if we know we are working in a path-connected space.
\end{boxconvention}

\begin{boxlemma}
    Let $X$ and $Y$ be (path-connected) spaces and let $f_0, f_1 : X \to Y$ be homotopic. Then, $\parenth{f_0}_* = \parenth{f_1}_*$.
\end{boxlemma}
\begin{proof}
    Let $H : X \times I \to Y$ be a homotopy between $f_0$ and $f_1$. The idea is to show that for all $\gamma \in \pi_1\of{X}$, the path $f_1 \circ \gamma$ is homotopic to the path $f_0 \circ \gamma$. \sorry
\end{proof}

% TODO: This should precede the discussion on deformation retractions, so it should eventually be moved. In fact all of this should be reorganised in to a new section.
\begin{boxtheorem}
    If we have $X\simeq Y$ then $\pi_1(X, x_0)\cong \pi_1(Y, y_0)$.
\end{boxtheorem}
\begin{proof}
    Assume we have (continuous) functions $f : X \to Y$ and $g : Y \to X$ such that $f \circ g \simeq \id_Y$ and $g \circ f \simeq \id_X$. We know show that $f_* : \pi_1\of{X, x_0} \to \pi_1\of{Y, f\of{x_0}}$ is a homomorphism; in fact, we will show it is an \textit{iso}morphism. It may not be true that $y_0 = f\of{x_0}$, but since $Y$ is path-connected, we know we have an isomorphism between $\pi_1\of{Y, f\of{x_0}}$ and $\pi_1\of{Y, y_0}$.

    \sorry % Just apply preceding lemma
\end{proof}


